---
title: "R practice"
author: "Frankie Fan"
date: "6/2/2022"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

## Intro to Markdown

Markdown allows you to do things like **bold**, *italic*, and more.

Make a list with asterisks like:

\* bold with double asterisks

\* italic with underscore

\* code-type fond with brackticks

You can also use hyphens instead of asterisks:

\- bold with double-asterisks

\- italics with underscores

\- code-type font with backticks

Be consistent!

For a numbered list:

1.  bold with double-asterisks
2.  italics with underscores
3.  code-type font with backticks


Hyperlinks are like [Hyperlinked text](https://mylink.org)

```{r load_data}
gap <- gapminder


```


```{r make-plot}
plot(lifeExp ~ year, data=gap)

```
```{r}
# this line will not run
temp <- c(1:3)  # this chunk of a line will not run



```
shortcut: option - for <-

```{r make_cats}
cats <- data.frame(coat = c("calico", "black", "tabby"),
                   weight = c(2.1, 5.0, 3.2),
                   likes_string = c(1, 0, 1))
write.csv(x=cats, file = "data/feline-data.csv", row.names = FALSE)
```

```{r read_cats}
cats2<- read.csv(file = "data/feline-data.csv", stringsAsFactors = TRUE)

```
paste("My cat is", cats$coat)

rm() removes something from the environment

typeof() shows you the data type;
cats$likes_string <- as.integer(cats$likes_string)

```{r}
names(cats)[3] <- "likes_hunting"
age <- c(2, 3, 5)
cats <- cbind(cats, age)

new_row <- list("tortoiseshell", 3.3, TRUE, 9)
cats <- rbind(cats, new_row)
```

Challenge
```{r}
human_age <- 7 * cats$age
human_age <- factor(human_age)
human_age <- as.numeric(as.character(human_age))
human_age / 7
```

Remove rows and columns
```{r}
cats$coat <- as.factor(cats$coat)
cats <- rbind(cats, c("white", 3.4, 0, 1))

# Removes the fifth row (similar for removing columns)
cats <- cats[-5, ]

# Removes all na values:
cats <- na.omit(cats)

drop <- names(cats) %in% 
  c("weight")
cats[, !drop]

```
shortcut: command-option-i for adding an r chunk

```{r}
gapminder <- read_excel("data/gapminder_data.xlsx")
summary(gapminder)

gapminder$country <- factor(gapminder$country)
gapminder$continent <- factor(gapminder$continent)

length(gapminder) #ncol gapminder
nrow(gapminder)
colnames(gapminder)

head(gapminder)
tail(gapminder)

gapminder[sample(nrow(gapminder), 10),]

```
## Subsetting Data
```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
names(x) <- c('a', 'b', 'c', 'd', 'e')

x[c(1, 3)]

x[-c(1, 5)]
x[c(-1, -5)] #the same thing

x[-(1:3)]

#by name
x[c("a", "c")]

x[x>7]

x[!names(x) %in% c("a", "c")]


```
Select Southeast Asian countries from gapminder

```{r}
seAsia <- c("Myanmar","Thailand","Cambodia","Vietnam","Laos")

seAsia_gap <- gapminder[gapminder$country %in% seAsia, ]

gapminder[c(seAsia)]
 
```

